---
title: "CedarWestRxDataAnalysis"
author: "Muscettola"
date: "2025-12-07"
output: html_document
---

#R setup

```{r}
#install packages
#install.packages(c("ggplot2", "dplyr", "tidyverse"))


#load libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(viridis)

#set working directory
path_data <- "S:/Protection/Fire/Ecology/effects/FireEffects_WorkingFolder/Captain Core Documents/Data_Analysis/Data/2025CedarWestRx/Raw/"
path_out <- "S:/Protection/Fire/Ecology/effects/FireEffects_WorkingFolder/Captain Core Documents/Data_Analysis/Data/2025CedarWestRx/R_working/"

#set working directory
path_data <- "S:/Protection/Fire/Ecology/effects/FireEffects_WorkingFolder/Captain Core Documents/Data_Analysis/Data/2025CedarWestRx/"

#load data 
FWD_raw <- read.csv(paste0(path_data, "Surface Fuels - Fine_XPT.csv"))
CWD_raw <-read.csv(paste0(path_data, "Surface Fuels - 1000Hr_XPT.csv"))
DuffLitt_raw <- read.csv(paste0(path_data, "Surface Fuels - Duff_Litter_XPT.csv"))
Trees_raw <-read.csv(paste0(path_data, "Trees - Individuals (metric)_XPT.csv"))
Seedlings_raw <- read.csv(paste0(path_data, "Density - Quadrats (metric)_XPT.csv"))
PostBurnSev_raw <- read.csv(paste0(path_data, "Post Burn Severity_XPT.csv"))
```

#Data Tidying
Note: Only do once. After the data is tidied, the code saves the tidy data as a csv so it can be quickly referenced in the future.
##12/7/2025 IM Consider removing non-pertinent columns (i.e. trees remove CrFuBHt:CKR)

###FWD
```{r}
#clean up dataframe 

#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
FWD <- FWD_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) %>%
  # set column types
  mutate(Plot=as.factor(Plot), Year = as.numeric(Year), Monitoring.Status = factor(Monitoring.Status, levels = c("Pre01", "Post01", "Re01Year01", "Re01Year02", "Re01Year05", "Re01Year10", "Pre02", "Post02", "Re02Year01", "Re02Year05", "Pre03", "Post03")))

#Ensure blanks in visited column are NA
FWD$Visited[FWD$Visited==""] <- NA

# Filter for only data
FWD_data <- FWD %>%
  filter(is.na(Visited))

#Filter for only headers
FWD_header <- FWD %>%
  filter(!is.na(Visited))

rm(FWD, FWD_raw)

```

### CWD
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
CWD <- CWD_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) %>%
  # set column types
  mutate(Plot=as.factor(Plot), Year = as.numeric(Year), Monitoring.Status = factor(Monitoring.Status, levels = c("Pre01", "Post01", "Re01Year01", "Re01Year02", "Re01Year05", "Re01Year10", "Pre02", "Post02", "Re02Year01", "Re02Year05", "Pre03", "Post03")))

#Ensure blanks in visited column are NA
CWD$Visited[CWD$Visited==""] <- NA

# Filter for only data
CWD_data <- CWD %>%
  filter(is.na(Visited))

#Filter for only headers
CWD_header <- CWD %>%
  filter(!is.na(Visited))

rm(CWD, CWD_raw)
```
### DuffLitt
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
DuffLitt <- DuffLitt_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) %>%
  # set column types
  mutate(Plot=as.factor(Plot), Year = as.numeric(Year), Monitoring.Status = factor(Monitoring.Status, levels = c("Pre01", "Post01", "Re01Year01", "Re01Year02", "Re01Year05", "Re01Year10", "Pre02", "Post02", "Re02Year01", "Re02Year05", "Pre03", "Post03")))

#Ensure blanks in visited column are NA
DuffLitt$Visited[DuffLitt$Visited==""] <- NA

# Filter for only data
DuffLitt_data <- DuffLitt %>%
  filter(is.na(Visited))

#Filter for only headers
DuffLitt_header <- DuffLitt %>%
  filter(!is.na(Visited))

rm(DuffLitt, DuffLitt_raw)
```
### PostBurnSev
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
PostBurnSev <- PostBurnSev_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) %>%
  # set column types
  mutate(Plot=as.factor(Plot), Year = as.numeric(Year), Monitoring.Status = factor(Monitoring.Status, levels = c("Pre01", "Post01", "Re01Year01", "Re01Year02", "Re01Year05", "Re01Year10", "Pre02", "Post02", "Re02Year01", "Re02Year05", "Pre03", "Post03")))

#Ensure blanks in visited column are NA
PostBurnSev$Visited[PostBurnSev$Visited==""] <- NA

# Filter for only data
PostBurnSev_data <- PostBurnSev %>%
  filter(is.na(Visited))

#Filter for only headers
PostBurnSev_header <- PostBurnSev %>%
  filter(!is.na(Visited))

rm(PostBurnSev, PostBurnSev_raw)
```
### Seedlings 
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
Seedlings <- Seedlings_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) %>%
  # set column types
  mutate(Plot=as.factor(Plot), Year = as.numeric(Year), Monitoring.Status = factor(Monitoring.Status, levels = c("Pre01", "Post01", "Re01Year01", "Re01Year02", "Re01Year05", "Re01Year10", "Pre02", "Post02", "Re02Year01", "Re02Year05", "Pre03", "Post03")))

#Ensure blanks in visited column are NA
Seedlings$Visited[Seedlings$Visited==""] <- NA

# Filter for only data
Seedlings_data <- Seedlings %>%
  filter(is.na(Visited))

#Filter for only headers
Seedlings_header <- Seedlings %>%
  filter(!is.na(Visited))

#good! looking at the header, the sampling area for all of the reads is 250m2

rm(Seedlings, Seedlings_raw)
```

### Trees 
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
Trees <- Trees_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) %>%
  # set column types
  mutate(Plot=as.factor(Plot), Year = as.numeric(Year), Monitoring.Status = factor(Monitoring.Status, levels = c("Pre01", "Post01", "Re01Year01", "Re01Year02", "Re01Year05", "Re01Year10", "Pre02", "Post02", "Re02Year01", "Re02Year05", "Pre03", "Post03")))

#Ensure blanks in visited column are NA
Trees$Visited[Trees$Visited==""] <- NA

# Filter for only data
Trees_data <- Trees %>%
  filter(is.na(Visited))

#Filter for only headers
Trees_header <- Trees %>%
  filter(!is.na(Visited))

rm(Trees, Trees_raw)
```

### Save tidy data as csv's 
```{r}
#save tidy data as csv
write.csv(FWD_data, paste0(path_out, "CGWestRx_FWD_data.csv"))
write.csv(FWD_header, paste0(path_out, "CGWestRx_FWD_header.csv"))
write.csv(CWD_data, paste0(path_out, "CGWestRx_CWD_data.csv"))
write.csv(CWD_header, paste0(path_out, "CGWestRx_CWD_header.csv"))
write.csv(DuffLitt_data, paste0(path_out, "CGWestRx_DuffLitt_data.csv"))
write.csv(DuffLitt_header, paste0(path_out, "CGWestRx_DuffLitt_header.csv"))
write.csv(PostBurnSev_data, paste0(path_out, "CGWestRx_PostBurnSev_data.csv"))
write.csv(PostBurnSev_header, paste0(path_out, "CGWestRx_PostBurnSev_header.csv"))
write.csv(Seedlings_data, paste0(path_out, "CGWestRx_Seedlings_data.csv"))
write.csv(Seedlings_header, paste0(path_out, "CGWestRx_Seedlings_header.csv"))
write.csv(Trees_data, paste0(path_out, "CGWestRx_Trees_data.csv"))
write.csv(Trees_header, paste0(path_out, "CGWestRx_Trees_header.csv"))
```

# Calculating Fuel Loading
12/7/2025 IM - waiting for Nick Kichas for code. Rfuels is not downloading due to newer version of R than allowed. 
## Using Rfuels

###Rfuels setup
```{r}
#install Rfuels package
install.packages("devtools")
library(devtools)
devtools::install_github('danfosterfire/Rfuels')
install.packages("Rfuels")
install.packages("Rtools")

#load Rfuels package
library(Rfuels)
```

###Reformat


# Fuels Data Visualizations

## FWD

```{r}
#summarize the FWD data by Monitoring Status for all size classes of fuel. This summarizes the minimum, maximum, mean, standard deviation, and standard error for One, Ten, and Hundred hour fuel classes.
FWD_summ <- FWD_data %>%
  group_by(Monitoring.Status) %>%
  reframe(n=n(),
            `Min.`=c(min(OneHr),min(TenHr), min(HunHr)),
            `Max.`=c(max(OneHr),max(TenHr),max(HunHr)),
            `Mean`=c(mean(OneHr),mean(TenHr),mean(HunHr)),
            `SD`=c(sd(OneHr),sd(TenHr),sd(HunHr)),
            `SE`=c(sd(OneHr)/sqrt(n),sd(TenHr)/sqrt(n),sd(HunHr)/sqrt(n)),
            'Size_class'=c("One", "Ten", "Hun"))
```

# Seedlings Data Visualizations

Calculate trees/ac of seedlings
```{r}
# Calculate trees/ac of seedlings knowing that the sampling area was 250m2
#250m2 = 0.0612 ac; 250m2 = 0.025ha
#total area for all 4 plots is 1000m2 (4*250m2)

Seedlings_data <- Seedlings_data %>%
  mutate(TPHa = Count/250*10000, Height_ft = Height * 3.28084, TPA = Count/250*4046.8564) %>%
  mutate(Height = as.factor(Height), Height_ft = as.factor(Height_ft))

Density_all_m <- Seedlings_data %>%
  group_by(Monitoring.Status, Height, Status) %>%
  summarise(Count = sum(Count),
            TPHa = Count/1000*10000) 
#In order to graph by height and status in the same graph we need to create a merged monitoring and live/dead column
#there are no dead seedlings pre burn so let's add a row that has 0 for the TPHa for Dead Pre03
temp <- tibble(Monitoring.Status = as.factor("Pre03"), Height = as.factor("0.15"), Status = "D", Count = 0, TPHa = 0, MonStat = as.factor("Pre03D"))
Density_all_m <- Density_all_m %>%
  mutate(MonStat = paste0(Monitoring.Status, Status)) %>%
  bind_rows(temp) %>%
  mutate(MonStat = factor(MonStat, c("Pre03L", "Pre03D", "Post03L", "Post03D")))


Density_all_ft <- Seedlings_data %>%
  group_by(Monitoring.Status, Height_ft) %>%
  summarise(Count = sum(Count),
            TPA = Count/1000*4046.8564)
#it looks like something weird happened in the counts for Re01Year02 and Re01Year05; very high counts for seedlings <0.15m
```

### Plot Seedling density by size class
```{r}
#metric
a <- ggplot(Density_all_m, aes(x=Monitoring.Status, y=TPHa, fill=Height)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(title = "Seedlings per Ha by Size Class", x = "Monitoring Status", y = "Seedlings/ha")
a

#now let's look only at the most recent burn
b <- ggplot(Density_all_m[36:48, ], aes(x=Monitoring.Status, y=TPHa, fill=Height)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(title = "Pre= and Post- Cedar West Rx: Seedlings per Ha by Size Class", x = "Monitoring Status", y = "Seedlings/ha")
b

#now looking only at the most recent burn, let's include the status of the seedlings (live or dead)
c <- ggplot(Density_all_m[49:66, ], aes(x=MonStat, y=TPHa, fill=Height)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(title = "Pre- and Post- Cedar West Rx: Seedlings per Ha by Size Class", x = "Monitoring Status Live/Dead", y = "Seedlings/ha") + geom_vline(xintercept = 2.5, linetype="dashed", color = "red", linewidth = 0.5)
c
```

###Plot Seedling size classes by species

make summary dataframes for bargraphs
```{r}
Density_sp_m <- Seedlings_data %>%
  group_by(Monitoring.Status, Height, Status, Species.Symbol) %>%
  summarise(Count = sum(Count),
            TPHa = Count/1000*10000) 

Density_all_ft <- Seedlings_data %>%
  group_by(Monitoring.Status, Height_ft, Status, Species.Symbol) %>%
  summarise(Count = sum(Count),
            TPA = Count/1000*4046.8564)
#it looks like something weird happened in the counts for Re01Year02 and Re01Year05; very high counts for seedlings <0.15m
=======
PostBurnSev_raw <- read.csv(paste0(path_data, "Post Burn Severity_XPT.csv"))
>>>>>>> feb29b897111fa709b6659a8362d27a712b06855
```

#Data Tidying
#12/7/2025 IM Consider removing non-pertinent columns (i.e. trees remove CrFuBHt:CKR)

###FWD
```{r}
#clean up dataframe 

#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
FWD <- FWD_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) 

#Ensure blanks in visited column are NA
FWD$Visited[FWD$Visited==""] <- NA

# Filter for only data
FWD_data <- FWD %>%
  filter(is.na(Visited))

#Filter for only headers
FWD_header <- FWD %>%
  filter(!is.na(Visited))

rm(FWD, FWD_raw)
```

### CWD
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
CWD <- CWD_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) 

#Ensure blanks in visited column are NA
CWD$Visited[CWD$Visited==""] <- NA

# Filter for only data
CWD_data <- CWD %>%
  filter(is.na(Visited))

#Filter for only headers
CWD_header <- CWD %>%
  filter(!is.na(Visited))

rm(CWD, CWD_raw)
```
### DuffLitt
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
DuffLitt <- DuffLitt_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) 

#Ensure blanks in visited column are NA
DuffLitt$Visited[DuffLitt$Visited==""] <- NA

# Filter for only data
DuffLitt_data <- DuffLitt %>%
  filter(is.na(Visited))

#Filter for only headers
DuffLitt_header <- DuffLitt %>%
  filter(!is.na(Visited))

rm(DuffLitt, DuffLitt_raw)
```
### PostBurnSev
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
PostBurnSev <- PostBurnSev_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) 

#Ensure blanks in visited column are NA
PostBurnSev$Visited[PostBurnSev$Visited==""] <- NA

# Filter for only data
PostBurnSev_data <- PostBurnSev %>%
  filter(is.na(Visited))

#Filter for only headers
PostBurnSev_header <- PostBurnSev %>%
  filter(!is.na(Visited))

rm(PostBurnSev, PostBurnSev_raw)
```
### Trees 
```{r}
#clean up date, extract year of the visit, separate MacroPlot.Name into Monitoring Type and Plot Number
Trees <- Trees_raw %>%
  #fix the naming convention of Cedar Plot 1
  mutate(MacroPlot.Name = ifelse(startsWith(MacroPlot.Name, "Cedar"), "B:FPIPO1T09:001", MacroPlot.Name)) %>%
  #Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  # Separate MacroPlot.Name to get Monitoring type
  separate(MacroPlot.Name, sep = ":", into = c("BurnType", "MonitoringType", "Plot"), remove = FALSE) %>%
  # Remove relic X column
  select(!X) 

#Ensure blanks in visited column are NA
Trees$Visited[Trees$Visited==""] <- NA

# Filter for only data
Trees_data <- Trees %>%
  filter(is.na(Visited))

#Filter for only headers
Trees_header <- Trees %>%
  filter(!is.na(Visited))

rm(Trees, Trees_raw)
```

```{r}
#dataframe of only Pre03 and Post03
seedlings_prepost03 <- filter(Seedlings_data, Monitoring.Status == "Pre03" | Monitoring.Status == "Post03")
Density_prepost03_m <- filter(Density_sp_m, Monitoring.Status=="Pre03" | Monitoring.Status == "Post03") 

#looking only at the most recent burn, let's include the status of the seedlings (live or dead)

#boxplot
d <- ggplot(Density_prepost03_m, aes(x=Monitoring.Status, y=TPHa, fill=Height)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Species.Symbol) +
  theme_bw() +
  labs(title = "Pre- and Post- Cedar West Rx: Seedlings per Ha by Size Class and Species", x = "Monitoring Status", y = "Seedlings/ha") + geom_vline(xintercept = 1.5, linetype="dashed", color = "red", linewidth = 0.5)
d


```


Size  Class density analysis
```{r}
#plot mean density for size class each year and total
ggplot(Seedlings_density, aes(x=Monitoring.Status, y=density_mean))
```



