---
title: "ShortReburnIntervals"
author: "Isabella Muscettola"
date: "2026-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Data and Libraries
```{r}
# install packages (if needed)
#install.packages("tidyverse", "Rtools", "dplyr", "knitr", "writexl", "ggplot2", "vegan", "patchwork")
#install.packages("patchwork")

#load packages
library(ggplot2)
library(tidyverse)
library(dplyr)
library(paletteer)
library(vegan)
library(patchwork)
library(stringr)
```

## Adjust File Paths 
```{r}
path_data <- "C:/Users/imuscettola/OneDrive - DOI/Captain_9/Data_Analysis/Data/ShrubShortReburn/"
path_figures <- "C:/Users/imuscettola/OneDrive - DOI/Captain_9/Data_Analysis/Figures/ShrubShortReburn/"
```


# ONLY DO ONCE: Set-up 

### Load Data
```{r}
ADFA_CoverPoints_RAW <- read.csv(paste0(path_data, "ADFA_CoverPoints.csv"))
ADFA_DensityBelts_RAW <- read.csv(paste0(path_data, "ADFA_DensityBelts.csv"))
ADFA_PostBurnSeverity_RAW <- read.csv(paste0(path_data, "ADFA_PostBurnSeverity.csv"))
ADFA_SpeciesComp_RAW <- read.csv(paste0(path_data, "ADFA_SpeciesComp.csv"))

CEBE_CoverPoints_RAW <- read.csv(paste0(path_data, "CEBE_CoverPoints.csv"))
CEBE_DensityBelts_RAW <- read.csv(paste0(path_data, "CEBE_DensityBelts.csv"))
CEBE_PostBurnSeverity_RAW <- read.csv(paste0(path_data, "CEBE_PostBurnSeverity.csv"))
CEBE_SpeciesComp_RAW <- read.csv(paste0(path_data, "CEBE_SpeciesComp.csv"))

SpList <- read.csv("C:/Users/imuscettola/OneDrive - DOI/Captain_9/Data_analysis/Data/20260203_SEKISpList.csv")

#merge data from different monitoring types by Protocol
CoverPoints_merge <- rbind(ADFA_CoverPoints_RAW, CEBE_CoverPoints_RAW)
CoverSpComp_merge <- rbind(ADFA_SpeciesComp_RAW, CEBE_SpeciesComp_RAW)
DensityBelts_merge <- rbind(ADFA_DensityBelts_RAW, CEBE_DensityBelts_RAW)
PostBurnSev_merge <- rbind(ADFA_PostBurnSeverity_RAW, CEBE_PostBurnSeverity_RAW)

rm(ADFA_CoverPoints_RAW, ADFA_DensityBelts_RAW, ADFA_PostBurnSeverity_RAW, ADFA_SpeciesComp_RAW,CEBE_CoverPoints_RAW, CEBE_DensityBelts_RAW, CEBE_PostBurnSeverity_RAW, CEBE_SpeciesComp_RAW)
```

## Clean up data
FFI exports data and metadata into a single csv file. We need to separate the data from the metadata and do other tidying.


###Species List
```{r}
#rename Species Symbol on Species list for subsequent dataframe merging & tidy
SpList <- SpList %>%
  rename(Species.Symbol=LocalSpecies_Symbol, Scientific.Name = MasterSpecies_ScientificName, Common.Name = LocalSpecies_CommonName, Family = MasterSpecies_Family, Genus = MasterSpecies_Genus, Prf..Lifeform = LocalSpecies_PreferedLifeForm, Life.Cycle = LocalSpecies_LifeCycle, Native = LocalSpecies_Nativity) %>%
  select(Species.Symbol, Scientific.Name, Common.Name, Family, Genus, Prf..Lifeform, Life.Cycle, Native) %>%
  mutate(Native = ifelse(Native == "True", "Native", "NonNative"))
```

### Point Intercept Vegetation Cover Protocol
```{r}
CoverPoints <- CoverPoints_merge %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  #separate macroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = ":", into = c("temp", "MonitoringType", "Plot"), remove = FALSE) %>%
  mutate(Treatment = ifelse(temp=="B", "Burn", "Control")) %>%
  #remove extraneous columns
  select(!c("CanopyLayer", "UV1", "UV2", "UV3", "Offset", "UV1Desc", "UV2Desc", "UV3Desc", "SaComment", "X", "temp")) %>%
  mutate(FireHistory = case_when(MacroPlot.Name %in% c("B:BADFA1D04:012", "B:BADFA1D04:022") ~ "LP+KNP",
                                 MacroPlot.Name %in% c("B:BADFA1D04:013", "B:BADFA1D04:014") ~ "LP+KNP+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:001", "B:BCEBE:002", "B:BCEBE:003") ~ "Castle",
                                 MacroPlot.Name %in% c("B:BCEBE:004", "B:BCEBE:005", "B:BCEBE:006") ~ "Castle+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:007", "B:BCEBE:008", "B:BCEBE:009") ~ "CoffeePot"))

# Ensure blanks in Visited column are NA
CoverPoints$Visited[CoverPoints$Visited==""] <- NA

# Filter for only data
CoverPoints_data <- CoverPoints %>%
  filter(is.na(Visited)) %>%
  #remove extraneous columns
  select(!c("Time", "Index", "FieldTeam", "EntryTeam", "Visited")) %>%
  #add information about the plant species
  merge(., SpList, by = "Species.Symbol")
  

# Filter for only headers
CoverPoints_header <- CoverPoints %>%
  filter(!is.na(Visited)) %>%
  #remove extraneous columns
  select(!c("Time", "Index", "Transect", "Point", "Tape", "Order", "Height", "Status", "Comment", "Species.Symbol"))
  

#make sure Year is numeric
CoverPoints_header$Year <- as.numeric(CoverPoints_header$Year)

rm(CoverPoints_merge, CoverPoints)
```

###Species Observed Protocol
```{r}
###Species Observed Protocol
CoverSpComp <- CoverSpComp_merge %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  #separate macroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = ":", into = c("temp", "MonitoringType", "Plot"), remove = FALSE) %>%
  mutate(Treatment = ifelse(temp=="B", "Burn", "Control")) %>%
  #remove extraneous columns
  select(!c("SizeCl", "AgeCl", "Cover", "Height", "UV1", "UV2", "UV3", "UV1Desc", "UV2Desc", "UV3Desc", "SaComment", "X", "MinCovLevel", "Area", "X",  "temp")) %>%
  mutate(FireHistory = case_when(MacroPlot.Name %in% c("B:BADFA1D04:012", "B:BADFA1D04:022") ~ "LP+KNP",
                                 MacroPlot.Name %in% c("B:BADFA1D04:013", "B:BADFA1D04:014") ~ "LP+KNP+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:001", "B:BCEBE:002", "B:BCEBE:003") ~ "Castle",
                                 MacroPlot.Name %in% c("B:BCEBE:004", "B:BCEBE:005", "B:BCEBE:006") ~ "Castle+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:007", "B:BCEBE:008", "B:BCEBE:009") ~ "CoffeePot"))

# Ensure blanks in Visited column are NA
CoverSpComp$Visited[CoverSpComp$Visited==""] <- NA

# Filter for only data
CoverSpComp_data <- CoverSpComp %>%
  filter(is.na(Visited)) %>%
  #remove extraneous columns
  select(!c("FieldTeam", "EntryTeam", "Visited", "Time")) %>%
  #add information about the plant species
  merge(., SpList, by = "Species.Symbol")

# Filter for only headers
CoverSpComp_header <- CoverSpComp %>%
  filter(!is.na(Visited)) %>%
  #remove extraneous columns
  select(!c("Time", "Index", "Status", "Comment", "Species.Symbol"))
  

rm(CoverSpComp, CoverSpComp_merge)
```

###Density Belts (Shrubs)
```{r}
DensityBelts <- DensityBelts_merge %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  #separate macroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = ":", into = c("temp", "MonitoringType", "Plot"), remove = FALSE) %>%
  mutate(Treatment = ifelse(temp=="B", "Burn", "Control")) %>%
  #remove extraneous columns
  select(!c("SizeCl", "SubFrac", "UV1", "UV2", "UV3", "UV1Desc", "UV2Desc",  "UV3Desc", "SaComment", "X")) %>%
  mutate(FireHistory = case_when(MacroPlot.Name %in% c("B:BADFA1D04:012", "B:BADFA1D04:022") ~ "LP+KNP",
                                 MacroPlot.Name %in% c("B:BADFA1D04:013", "B:BADFA1D04:014") ~ "LP+KNP+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:001", "B:BCEBE:002", "B:BCEBE:003") ~ "Castle",
                                 MacroPlot.Name %in% c("B:BCEBE:004", "B:BCEBE:005", "B:BCEBE:006") ~ "Castle+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:007", "B:BCEBE:008", "B:BCEBE:009") ~ "CoffeePot"))

# Ensure blanks in Visited column are NA
DensityBelts$Visited[DensityBelts$Visited==""] <- NA

# Filter for only data
DensityBelts_data <- DensityBelts %>%
  filter(is.na(Visited)) %>%
  #make sure area is correct - this should be double checked for each data set before setting
  mutate(TranLen = 30, TranWid = 1, Area = 30) %>% 
  #remove unnecessary columns
  select(!c("Index", "Height", "NumTran", "NumSubbelt", "FieldTeam", "EntryTeam", "Visited")) %>%
  #add information about the plant species
  merge(., SpList, by = "Species.Symbol")

# Filter for only headers
DensityBelts_header <- DensityBelts %>%
  filter(!is.na(Visited)) %>%
  #make sure NumSubbelt is correct - double check for your data set before setting
  mutate(NumSubbelt = 6) %>%
  #remove unnecessary columns
  select(!c("Index", "Transect", "Subbelt", "Status", "AgeCl", "Count", "Height", "Species.Symbol"))

rm(DensityBelts_merge, DensityBelts)
```

###Post Burn Severity
```{r}
PostBurnSev <- PostBurnSev_merge %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day")) %>%
  #separate macroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = ":", into = c("temp", "MonitoringType", "Plot"), remove = FALSE) %>%
  mutate(Treatment = ifelse(temp=="B", "Burn", "Control")) %>%
  #remove extraneous columns
  select(!c("UV1", "UV2", "UV3", "PlotType", "UV1Desc", "UV2Desc",  "UV3Desc", "SaComment", "X")) %>%
  mutate(FireHistory = case_when(MacroPlot.Name %in% c("B:BADFA1D04:012", "B:BADFA1D04:022") ~ "LP+KNP",
                                 MacroPlot.Name %in% c("B:BADFA1D04:013", "B:BADFA1D04:014") ~ "LP+KNP+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:001", "B:BCEBE:002", "B:BCEBE:003") ~ "Castle",
                                 MacroPlot.Name %in% c("B:BCEBE:004", "B:BCEBE:005", "B:BCEBE:006") ~ "Castle+CoffeePot",
                                 MacroPlot.Name %in% c("B:BCEBE:007", "B:BCEBE:008", "B:BCEBE:009") ~ "CoffeePot"))

# Ensure blanks in Visited column are NA
PostBurnSev$Visited[PostBurnSev$Visited==""] <- NA

# Filter for only data
PostBurnSev_data <- PostBurnSev %>%
  filter(is.na(Visited)) %>%
  #remove extraneous columns
  select(!c("Time", "NumTran", "TranLen", "NumPtsTran","PtArea", "FieldTeam", "EntryTeam", "Visited"))

# Filter for only headers
PostBurnSev_header <- PostBurnSev %>%
  filter(!is.na(Visited)) %>%
  #remove extraneous columns
  select(!c("Time", "Index", "Transect", "Point", "TapeDist", "Veg", "Sub", "Comment"))

rm(PostBurnSev_merge, PostBurnSev)
```


### Save merged and tidied dataframes
```{r}
#write.csv(CoverSpComp_data, paste0(path_data, "2_working/SpComp_data.csv"))
#write.csv(CoverSpComp_header, paste0(path_data, "2_working/SpComp_header.csv"))
#write.csv(CoverPoints_data, paste0(path_data, "2_working/Points_data.csv"))
#write.csv(CoverPoints_header, paste0(path_data, "2_working/Points_header.csv"))
#write.csv(DensityBelts_data, paste0(path_data, "2_working/DensityBelts_data.csv"))
#write.csv(DensityBelts_header, paste0(path_data, "2_working/DensityBelts_header.csv"))
#write.csv(PostBurnSev_data, paste0(path_data, "2_working/PostBurnSev_data.csv"))
#write.csv(PostBurnSev_header, paste0(path_data, "2_working/PostBurnSev_header.csv"))
#write.csv(SpList, paste0(path_data, "2_working/SpList.csv"))
```

## Data summaries

### Unique Species List for ADFA and CEBE plots
```{r}
UniquePoints <- CoverPoints_data %>%
  group_by(MonitoringType) %>%
  distinct(Species.Symbol) %>%
  #add species metadata
  merge(., SpList, by = "Species.Symbol")

UniqueSpComp <- CoverSpComp_data %>%
  group_by(MonitoringType) %>%
  distinct(Species.Symbol) %>%
  #add species metadata
  merge(., SpList, by = "Species.Symbol")

ADFA_UniqueSpecies <- rbind(UniquePoints, UniqueSpComp) %>%
  filter(MonitoringType == "BADFA1D04") %>%
  distinct(., .keep_all=TRUE) %>%
  #move the Monitoring Type column to the beginning of the table
  relocate(MonitoringType, .before = 'Species.Symbol')

CEBE_UniqueSpecies <- rbind(UniquePoints, UniqueSpComp) %>%
  filter(MonitoringType == "BCEBE") %>%
  distinct(., .keep_all=TRUE) %>%
  #move the Monitoring Type column to the beginning of the table
  relocate(MonitoringType, .before = 'Species.Symbol')

UniqueSpecies <- rbind(UniquePoints, UniqueSpComp) %>%
  distinct(., .keep_all=TRUE) %>%
  #remove extraneous column
  select(!MonitoringType)

#now lets do it by plot and read
temp <-  CoverPoints_data %>%
  group_by(MacroPlot.Name, Monitoring.Status, Year, FireHistory) %>%
  distinct(Species.Symbol) %>%
  #add species metadata
  merge(., SpList, by = "Species.Symbol")

temp2 <- SpComp_data %>%
  group_by(MacroPlot.Name, Monitoring.Status, Year, FireHistory) %>%
  distinct(Species.Symbol) %>%
  #add species metadata
  merge(., SpList, by = "Species.Symbol")

UniqueSpecies_byPlotRead <- rbind(temp, temp2) %>%
  group_by(MacroPlot.Name, Monitoring.Status, Year, FireHistory) %>%
  distinct(Species.Symbol) %>%
  merge(., SpList, by = "Species.Symbol") %>%
  relocate(MacroPlot.Name:Year, .before = 'Species.Symbol') %>%
  relocate(FireHistory, .before = 'Species.Symbol')

#save species lists to OneDrive
#write.csv(ADFA_UniqueSpecies, paste0(path_data, "3_final/ADFA_LookoutPoint_UniqueSpecies.csv"))
#write.csv(CEBE_UniqueSpecies, paste0(path_data, "3_final/CEBE_UniqueSpecies.csv"))
#write.csv(UniqueSpecies, paste0(path_data, "3_final/ShrubShortReburn_UniqueSpecies.csv"))
#write.csv(UniqueSpecies_byPlotRead, paste0(path_data, "2_working/UniqueSpecies_byPlotRead.csv"))

rm(UniqueSpComp, UniquePoints, UniqueSpecies, ADFA_UniqueSpecies, CEBE_UniqueSpecies, temp, temp2, UniqueSpecies_byPlotRead)
```

### Summarize data
Summarize the Point Intercept Data so that we can visualize species by abundance (% of times that species was encountered in each read)

```{r}
#count total number of hits per read
temp <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status) %>%
  summarise(total.hits=n())

#count total number of hits of each species per read
temp2 <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status, Species.Symbol) %>%
  summarise(sp.hits=n())

#merge summary calculations into one dataframe including species information
CoverPoints_summary <- merge(temp, temp2, by=c("Plot", "Monitoring.Status")) %>%
  #include species information
  merge(., SpList, by="Species.Symbol") %>%
  #include header information
  merge(CoverPoints_header, ., by=c("Plot", "Monitoring.Status")) %>%
  #calculate the % of hits for each species
  mutate(., perc_of_hits = sp.hits/total.hits*100) 

#remove temporary dataframes
rm(temp, temp2)

#save dataframe to S drive
#write.csv(CoverPoints_summary, paste0(path_data, "2_working/Points_summarybyspecies.csv"))

```

Summarize the Point Intercept Data - by Species and Status (ADFA Dead vs. ADFA Live)

```{r}
#count total number of hits per read
temp <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status) %>%
  summarise(total.hits=n())

#count total number of hits of each species per read
temp2 <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status, Species.Symbol, Status) %>%
  summarise(sp.hits=n())

#merge summary calculations into one dataframe including species information
CoverPoints_summary <- merge(temp, temp2, by=c("Plot", "Monitoring.Status")) %>%
  #include species information
  merge(., SpList, by="Species.Symbol") %>%
  #include header information
  merge(CoverPoints_header, ., by=c("Plot", "Monitoring.Status")) %>%
  #calculate the % of hits for each species
  mutate(., perc_of_hits = sp.hits/total.hits*100) 

#remove temporary dataframes
rm(temp, temp2)

#save dataframe to S drive
#write.csv(CoverPoints_summary, paste0(path_data, "2_working/Points_summarybyspeciesandstatus.csv"))

```

Now let's summarize the Point Intercept data according to preferred life form so that we can visualize species by abundance (% of times that species was encountered in each read)

```{r}
#count the number of hits of each lifeform for all plots and reads
temp3 <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status, Prf..Lifeform) %>%
  summarise(lifeform.hits=n())

#count total number of hits per read
temp <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status) %>%
  summarise(total.hits=n())

#add the metadata from the header dataframe
Points_sum_lifeform <- merge(CoverPoints_header, temp3, by=c("Plot", "Monitoring.Status")) %>%
  # add the column of total hits for each read
  merge(., temp, by=c("Plot", "Monitoring.Status")) %>%
  # calculate the % of hits for each species
  mutate(., perc_of_hits =lifeform.hits/total.hits*100)

#remove temporary dataframes
rm(temp, temp3)

#save Lifeform summary dataframe
#write.csv(Points_sum_lifeform, paste0(path_data, "2_working/Points_summarybylifeform.csv"))
```

Summarize by lifeform and status
```{r}
#count the number of hits of each lifeform for all plots and reads
temp3 <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status, Prf..Lifeform, Status) %>%
  summarise(lifeform.hits=n())

#count total number of hits per read
temp <- CoverPoints_data %>%
  group_by(Plot, Monitoring.Status) %>%
  summarise(total.hits=n())

#add the metadata from the header dataframe
Points_sum_lifeform <- merge(CoverPoints_header, temp3, by=c("Plot", "Monitoring.Status")) %>%
  # add the column of total hits for each read
  merge(., temp, by=c("Plot", "Monitoring.Status")) %>%
  # calculate the % of hits for each species
  mutate(., perc_of_hits =lifeform.hits/total.hits*100)

#remove temporary dataframes
rm(temp, temp3)

#save Lifeform summary dataframe
write.csv(Points_sum_lifeform, paste0(path_data, "2_working/Points_summarybylifeformandstatus.csv"))
```


### Remove all data it will be loaded for each analyses from the file directory
```{r}
rm(CoverPoints_data, CoverPoints_header, CoverSpComp_data, CoverSpComp_header, DensityBelts_data, DensityBelts_header, PostBurnSev_data, PostBurnSev_header, SpComp_data, SpComp_header, SpList, CoverPoints_summary, Points_sum_lifeform)
```

# Start Here!

## Load Tidied Data
```{r}
CoverPoints_data <- read.csv(paste0(path_data, "2_working/Points_data.csv"))
CoverPoints_header <- read.csv(paste0(path_data, "2_working/Points_header.csv"))
SpComp_data <- read.csv(paste0(path_data, "2_working/SpComp_data.csv"))
SpComp_header <- read.csv(paste0(path_data, "2_working/SpComp_header.csv"))
DensityBelts_data <- read.csv(paste0(path_data, "2_working/DensityBelts_data.csv"))
DensityBelts_header <- read.csv(paste0(path_data, "2_working/DensityBelts_header.csv"))
PostBurnSev_data <- read.csv(paste0(path_data, "2_working/PostBurnSev_data.csv"))
PostBurnSev_header <- read.csv(paste0(path_data, "2_working/PostBurnSev_header.csv"))
SpList <- read.csv(paste0(path_data, "2_working/SpList.csv"))
Points_summarybyspecies <- read.csv(paste0(path_data, "2_working/Points_summarybyspecies.csv"))
Points_summarybylifeform <- read.csv(paste0(path_data, "2_working/Points_summarybylifeform.csv"))
SpRichness_byPlotRead <- read.csv(paste0(path_data, "2_working/UniqueSpecies_byPlotRead.csv"))
Points_summarybyspeciesandstatus <- read.csv(paste0(path_data, "2_working/Points_summarybyspeciesandstatus.csv"))
Points_summarybylifeformandstatus <- read.csv(paste0(path_data, "2_working/Points_summarybylifeformandstatus.csv"))
```


## Initial Data Exploration

Species abundance per plot per read

```{r}
#change Year to factor so it graphs nicer
Points_summarybyspecies$Year <- as.factor(Points_summarybyspecies$Year)

ggplot(Points_summarybyspecies, aes(x=Year, y=perc_of_hits, fill=Species.Symbol), fill=Species.Symbol) +
  geom_col(position = "fill") + 
  facet_wrap(~MacroPlot.Name) +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=14), axis.title.x=element_text(size=20), axis.title.y=element_text(size=20),legend.title=element_text(size=14)) +
  labs(y="% Cover", fill="Species Symbol")
  

#save the figure to the S drive
#ggsave(paste0(path_figures, "SpeciesAbundancebyPlotandYear.png"), width = 14, height = 10, limitsize=FALSE)
```

There are so many species of plants it's hard to digest! Let's group and summarize by lifeform (i.e. shrub, tree, graminoid, etc.)

```{r}
Points_summarybylifeform$Year <- as.factor(Points_summarybylifeform$Year)

ggplot(Points_summarybylifeform, aes(x=Year, y=perc_of_hits, fill=Prf..Lifeform), fill=Prf..Lifeform) +
  geom_col(position = "fill") + 
  facet_wrap(~MacroPlot.Name) +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=14), axis.title.x=element_text(size=20), axis.title.y=element_text(size=20),legend.title=element_text(size=20), legend.text=element_text(size=14)) +
  labs(y="% Cover", fill="Cover Class") +
  scale_fill_paletteer_d("PrettyCols::Autumn")

#save the figure to the S drive
#ggsave(paste0(path_figures, "LifeformAbundancebyPlotandYear.png"), width = 14, height = 10, limitsize=FALSE)
```

## Tracking Shrub Abundance Across Time 

How has the abundance of shrubs changed across time, particularly with the occurrence of fire? 

NOTE: Fire Histories for each plot:
ADFA 12 & 22: 1998 (before 1998 plot read), 2021 (KNP)
ADFA 13 & 14: 1998, 2021 (KNP), 2024 (CoffeePot, before 2024 plot read)
CEBE 1-3: 2020 (Castle) only
CEBE 4-6: 2020 (Caste) + 2024 (CoffeePot)
CEBE 7-9: 2024 (CoffeePot) only

### Shrub % Cover across time
#### CEBE
```{r}
#Abundance of shrubs across time (Boxplot for CEBE plots & geom_point for ADFA plots)

#make sure Year is numeric
Points_summarybylifeform$Year <- as.numeric(Points_summarybylifeform$Year)

#boxplot
CEBE_summarybylifeform <- Points_summarybylifeform %>%
  filter(MonitoringType=="BCEBE") %>%
  mutate(Monitoring.Status = case_when(Monitoring.Status %in% c("Re01Year05", "Re01Year5") ~ "Re01Year05",
                                       Monitoring.Status=="Re01Year01" ~ "Re01Year01",
                                       Monitoring.Status=="Re02Year01" ~ "Re02Year01"))

ggplot(CEBE_summarybylifeform[CEBE_summarybylifeform$Prf..Lifeform=="Shrub", ], aes(x=Monitoring.Status, y=perc_of_hits, fill=FireHistory)) +
  geom_boxplot() + 
  theme_bw() + 
  labs(title='CEBE Plots - Shrub Cover', x='Monitoring Status', y='% Cover') +
  scale_x_discrete(limits= c("Re01Year05", "Re01Year01", "Re02Year01")) + 
  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20),legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold"))

#ggsave(paste0(path_figures, "CEBE_ShrubAbundancebyFireHistory_BoxPlot.png"), width = 8.5, height = 7, limitsize=FALSE)

```

```{r}
# looking at live vs dead shrub cover
CEBE_summarybylifeformandstatus <- Points_summarybylifeformandstatus %>%
  filter(MonitoringType=="BCEBE") %>%
  mutate(Monitoring.Status = case_when(Monitoring.Status %in% c("Re01Year05", "Re01Year5") ~ "Re01Year05",
                                       Monitoring.Status=="Re01Year01" ~ "Re01Year01",
                                       Monitoring.Status=="Re02Year01" ~ "Re02Year01"))
#

ggplot(CEBE_summarybylifeformandstatus[CEBE_summarybylifeformandstatus$Prf..Lifeform=="Shrub", ], aes(x=FireHistory, y=perc_of_hits, fill=Status)) +
         geom_boxplot() +
         theme_bw() +
         labs(title="CEBE Plots - 2025 Shrub Cover", x="Fire History", y = "% Cover") +
         scale_x_discrete(limits= c("Castle", "CoffeePot", "Castle+CoffeePot")) + 
  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20),legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold")) +
  scale_fill_manual(values = c("D" = "#707070", "L" = "#008B00"), labels=c("D"="Dead", "L"="Live"))

#ggsave(paste0(path_figures, "CEBE_ShrubAbundancebyFireHistoryandStatus_BoxPlot.png"), width = 8.5, height = 7, limitsize=FALSE)
```


#### Lookout Point ADFA - needs to be fixed
```{r}
#HOLD - missing ADFA data from 2022 plot read - geom_point
#ADFA_summarybylifeform <- Points_summarybylifeform %>%
#  filter(MonitoringType =="BADFA1D04")
#ADFA_summarybylifeform$Year <- as.numeric(ADFA_summarybylifeform$Year)

#ggplot(ADFA_summarybylifeform[ADFA_summarybylifeform$Prf..Lifeform=="Shrub", ], aes(x=Year, y=perc_of_hits, color=FireHistory, group=MacroPlot.Name)) +
#  geom_point() + 
#  geom_line() +
#  geom_vline(xintercept = c(1997, 2021, 2023.9), linetype = "dashed", color="red") +
#  theme_bw() + 
#  labs(title='ADFA Plots - Shrub Cover', x='Monitoring Status', y='% Shrub Cover') + 
#  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20),legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold"))

#ggsave(paste0(path_figures, "ADFA_ShrubAbundancebyFireHistory_PointsPlot.png"), width = 8.5, height = 7, limitsize=FALSE)

```

### % Shrub Cover by Species
#### CEBE Plots
```{r}
CEBE_summarybyspecies <- Points_summarybyspecies %>%
  filter(MonitoringType=="BCEBE") %>%
  mutate(Monitoring.Status = case_when(Monitoring.Status %in% c("Re01Year05", "Re01Year5") ~ "Re01Year05",
                                       Monitoring.Status=="Re01Year01" ~ "Re01Year01",
                                       Monitoring.Status=="Re02Year01" ~ "Re02Year01")) %>%
  mutate(FireHistory = case_when(FireHistory=="Castle" ~ "Castle",
                                 FireHistory=="CoffeePot" ~ "CoffeePot",
                                 FireHistory=="Castle+CoffeePot"~"Castle and CoffeePot"))


class(Points_summarybyspecies$FireHistory)
#Bar Plot showing the abundance of Shrubs across time colored by Species and faceted by treatment

#make sure year is numeric
CEBE_summarybyspecies$Year <- as.numeric(CEBE_summarybyspecies$Year)

ggplot(CEBE_summarybyspecies[CEBE_summarybyspecies$Prf..Lifeform=="Shrub", ], aes(x=FireHistory, y=perc_of_hits, fill=Species.Symbol)) + 
  geom_boxplot() + 
  facet_wrap(~Species.Symbol) +
  labs(title="CEBE Plots - Shrub Abundance by Species", x="Fire History", y="% Cover") +
  scale_x_discrete(limits=c("Castle", "CoffeePot", "Castle and CoffeePot"), labels = function(x) gsub(" ", "\n", x)) + 
  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20), legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold"))

#ggsave(paste0(path_figures, "CEBE_ShrubAbundance_bySpecies_BoxPlot.png"), width = 11, height=9, limitsize=FALSE)

#Let's just look at CEBE (mountain mahogany occurence)- the most common shrub and found in all 3 fire histories
ggplot(CEBE_summarybyspecies[CEBE_summarybyspecies$Species.Symbol=="CEBE1", ], aes(x=FireHistory, y=perc_of_hits)) +
  geom_boxplot() + 
  theme_bw() +
  labs(title="CEBE Plots - Mountain Mahogany Abundance", x="Fire History", y="% Cover") +
  scale_x_discrete(limits=c("Castle", "CoffeePot", "Castle and CoffeePot"), labels = function(x) gsub(" ", "\n", x)) + 
  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20), legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold"))

#ggsave(paste0(path_figures, "CEBE_CEBEAbundance_BoxPlot.png"), width = 8, height=6, limitsize=FALSE)
```

### Shrub Cover by Species and Status

#### CEBE
```{r}
CEBE_summarybyspeciesandstatus <- Points_summarybyspeciesandstatus %>%
  filter(MonitoringType=="BCEBE") %>%
  mutate(Monitoring.Status = case_when(Monitoring.Status %in% c("Re01Year05", "Re01Year5") ~ "Re01Year05",
                                       Monitoring.Status=="Re01Year01" ~ "Re01Year01",
                                       Monitoring.Status=="Re02Year01" ~ "Re02Year01")) %>%
  mutate(FireHistory = case_when(FireHistory=="Castle" ~ "Castle",
                                 FireHistory=="CoffeePot" ~ "CoffeePot",
                                 FireHistory=="Castle+CoffeePot"~"Castle and CoffeePot"))



#Bar Plot showing the abundance of Shrubs across time colored by Species and faceted by treatment

#make sure year is numeric
CEBE_summarybyspeciesandstatus$Year <- as.numeric(CEBE_summarybyspeciesandstatus$Year)

ggplot(CEBE_summarybyspeciesandstatus[CEBE_summarybyspeciesandstatus$Prf..Lifeform=="Shrub", ], aes(x=FireHistory, y=perc_of_hits, fill=Status)) + 
  geom_boxplot() + 
  facet_wrap(~Species.Symbol) +
  labs(title="CEBE Plots - Shrub Abundance by Species", x="Fire History", y="% Cover") +
  scale_x_discrete(limits=c("Castle", "CoffeePot", "Castle and CoffeePot"), labels = function(x) gsub(" ", "\n", x)) + 
  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20), legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold")) +
  scale_fill_manual(values = c("D" = "#707070", "L" = "#008B00"), labels=c("D"="Dead", "L"="Live"))

#ggsave(paste0(path_figures, "CEBE_ShrubAbundance_bySpeciesandStatus_BoxPlot.png"), width = 11, height=9, limitsize=FALSE)

#Let's just look at CEBE (mountain mahogany occurence)- the most common shrub and found in all 3 fire histories
ggplot(CEBE_summarybyspeciesandstatus[CEBE_summarybyspeciesandstatus$Species.Symbol=="CEBE1", ], aes(x=FireHistory, y=perc_of_hits, fill=Status)) +
  geom_boxplot() + 
  theme_bw() +
  labs(title="CEBE Plots - Mountain Mahogany Abundance", x="Fire History", y="% Cover") +
  scale_x_discrete(limits=c("Castle", "CoffeePot", "Castle and CoffeePot"), labels = function(x) gsub(" ", "\n", x)) + 
  theme(axis.text.x=element_text(size=14), axis.title.x=element_text(size=20), axis.text.y=element_text(size=14), axis.title.y=element_text(size=20), legend.title=element_text(size=20), legend.text=element_text(size=14), plot.title=element_text(size=20, face="bold")) +
  scale_fill_manual(values = c("D" = "#707070", "L" = "#008B00"), labels=c("D"="Dead", "L"="Live"))

#ggsave(paste0(path_figures, "CEBE_CEBEAbundancebyStatus_BoxPlot.png"), width = 8, height=6, limitsize=FALSE)
```

### Height of Shrubs colored by living or dead by fire history
```{r}

```

